Reproducing anaylses of Beninca et al (2008)

### 2016-07-03 article/article.tex

The article stated: ``For short-term forecasts of only a few days, most
species had a high predictability of R2 = 0.70 -- 0.90 (Fig. 2).
However, the predictability of the species was much reduced when
prediction times were extended to 15--30days.'' The reproduced
predictabilities, which were calculated from the GAMs, were consistent
with these qualitative statements, and were most often quantitatively
similar (\ref{fig:prediction_distance}). We did not reproduce the
predictability estimates for linear models.

@@

This reproduction was made as part of the Reproducible Research in
Ecology, Evolution, Behaviour, and Environmental Studies (RREEBES)
Course, lead by Owen Petchey at the University of Zurich. More
information about the course
\href{https://github.com/opetchey/RREEBES/blob/master/README.md}{here}
on github. Many thanks to Elisa Benincà and Stephen Ellner for
assistance, and to Stephen Ellner for providing unpublished data and
code.

>>>>

The article stated: ``For short-term forecasts of only a few days, most
species had a high predictability of R2 = 0.70 -- 0.90 (Fig. 2).
However, the predictability of the species was much reduced when
prediction times were extended to 15--30 days.'' The reproduced
predictabilities, which were calculated from the GAMs, were consistent
with these qualitative statements, and were most often quantitatively
similar (\ref{fig:prediction_distance}). We did not reproduce the
predictability estimates for linear models.

@@

This reproduction was made as part of the Reproducible Research in
Ecology, Evolution, Behaviour, and Environmental Studies (RREEBES)
Course, lead by Owen Petchey at the University of Zurich. More
information about the course
\href{https://github.com/opetchey/RREEBES/blob/master/README.md}{here}
on github. Many thanks to Elisa Benincà, Stephen Ellner and Jef Huisman
for assistance, and to Stephen Ellner for providing unpublished data and
code.

### 2016-07-03 code/report.Rmd

The exact method for identifying the "long sequences of zeros" was not given. However, the transformed data with zeros removed was provided, and from this we can get the day range of the retained data:
```{r}
tr <- read.csv(text=getURL(paste0(repo,"/data/original/transformed_data_Nature2008.csv")),
               skip=7, na.string="")
tr <- tr[,-14:-24] ## remove bad columns
tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
tr <- gather(tr, key="Species", value="Abundance", 2:13, convert=F) ## make long format
#tr$Species <- as.factor(tr$Species)
levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods" # rename a level (and next line)
levels(tr$Species)[levels(tr$Species)=="Total.Dissolved.Inorganic.Nitrogen"] <- "Total.dissolved.inorganic.nitrogen"
names(tr)[2] <- "variable" # rename a couple of variables...
names(tr)[1] <- "Day.number"
tr$Day.number <- as.numeric(as.character(tr$Day.number)) # make Day.number numeric
## get day range, in order to use this to constrain later data
min.max.day <- group_by(tr, variable) %>%
  summarise(min.day=min(Day.number[!is.na(Abundance)]), max.day=max(Day.number[!is.na(Abundance)]))
```

>>>>

The exact method for identifying the "long sequences of zeros" was not given. However, the transformed data with zeros removed was provided, and from this we can get the day range of the retained data:
```{r}
tr <- read.csv(text=getURL(paste0(repo,"/data/original/transformed_data_Nature2008.csv")),
               skip=7, na.string="")
tr <- tr[,-14:-24] ## remove bad columns
tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
tr <- gather(tr, key="Species", value="Abundance", 2:13, convert=F) ## make long format
tr$Species <- as.factor(tr$Species)
levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods" # rename a level (and next line)
levels(tr$Species)[levels(tr$Species)=="Total.Dissolved.Inorganic.Nitrogen"] <- "Total.dissolved.inorganic.nitrogen"
names(tr)[2] <- "variable" # rename a couple of variables...
names(tr)[1] <- "Day.number"
tr$Day.number <- as.numeric(as.character(tr$Day.number)) # make Day.number numeric
## get day range, in order to use this to constrain later data
min.max.day <- group_by(tr, variable) %>%
  summarise(min.day=min(Day.number[!is.na(Abundance)]), max.day=max(Day.number[!is.na(Abundance)]))
```
